si: code and prose exploring Spatial Interaction models

### 2022-05-15 README.qmd
Document bipartite SIMs, close #17

@@

The output `si_calculate)` is a geographic object which can be plotted as a map:

@@

```{r}
names(od)
```

As shown in the output above, the function allows you to use any variable in the origin or destination data in the function, with column names appended with `origin` and `destination`.

The approach works equally well when origin and destination points are different, as shown in the following example which estimates the number of trips to pubs in Leeds, assuming that for every trip to the pub there are 50 trips to work:

```{r pubs}
zones_pubs = si_zones %>% 
  mutate(to_pubs = all / 50)
pubs_example = si_pubs %>% 
  filter(grepl(pattern = "Chemic|Nag", x = name))
pubs_example$size = c(100, 80)
od_to_pubs = si_to_od(zones_pubs, pubs_example)
od_to_pubs_result = od_to_pubs %>% 
  si_calculate(fun = gravity_model, 
               m = origin_to_pubs,
               n = destination_size,
               d = distance_euclidean,
               beta = 0.5,
               constraint_production = origin_to_pubs)
od_to_pubs_result %>% 
  select(O, D, destination_name, interaction)
```

@@

```{r pubmap}
library(tmap)
tmap_mode("view")
od_to_pubs_result %>% 
  top_n(n = 50, wt = interaction) %>% 
  tm_shape() +
  tm_lines(col = "interaction", palette = "viridis", scale = 2)
```

>>>>

Unlike many software tools designed to support spatial interaction modelling, {si} does not define (or even encourage use of) any particular functional forms or modelling frameworks for predicting movement between origins and destinations.
Instead, it provides a framework enabling you to use model function forms or models of your choosing, ensuring flexibility and encouraging flexibility.

@@

The output of `si_calculate()` is a geographic object that can be plotted with `sf`'s plot method (or other geographic data visualisation packages):

@@

As documented in the [si vignette](https://robinlovelace.github.io/si/articles/si.html), the function allows you to use any variable in the origin or destination data by joining all attributes onto the OD data frame, with column names appended with `origin` and `destination`.

The approach works equally well for 'bipartite' SIMs, in which origin and destination points are different (Hasova et al. [2022](https://lenkahas.com/files/preprint.pdf)).
The following example implements a bipartite SIM that estimates the number of trips from administrative zones to pubs in Leeds:

```{r pubs}
# Set n. trips to pubs, assuming that for every trip to the pub there are
# 50 trips to work (this would be validated/tested/modelled in empirical work)
zones_pubs = si_zones %>% 
  mutate(to_pubs = all / 50)
pubs_example = si_pubs %>% 
  filter(grepl(pattern = "Chemic|Nag", x = name))
pubs_example$size = c(100, 80)
od_to_pubs = si_to_od(zones_pubs, pubs_example)
od_to_pubs_result = od_to_pubs %>% 
  si_calculate(fun = gravity_model, 
               m = origin_to_pubs,
               n = destination_size,
               d = distance_euclidean,
               beta = 0.5,
               constraint_production = origin_to_pubs)
od_to_pubs_result %>% 
  select(O, D, destination_name, interaction)
```

@@

```{r pubmap, eval=FALSE}
library(tmap)
tmap_mode("view")
od_to_pubs_result %>% 
  top_n(n = 50, wt = interaction) %>% 
  tm_shape() +
  tm_lines(col = "interaction", palette = "viridis", scale = 2)
```

![](https://user-images.githubusercontent.com/1825120/168485930-1adf9970-7dc1-4417-891e-2d5a1ec5e170.png)
